A multi-dimensional extension of the Lightweight Temporal Compression method

### 2018-10-09 paper.tex
Finished section II, advanced in section III

% packages
\usepackage{algorithm}
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%\usepackage{algpseudocode}
\usepackage{amsfonts} % for R symbol (the set of real numbers)
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[bookmarks=false]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,filecolor=black,urlcolor=blue}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{subcaption}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amsthm}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\title{A multi-dimensional extension of the Lightweight Temporary Compression method}

@@

We use the following notations and assumptions to describe the LTC algorithm.
The algorithm \emph{receives} a stream of \emph{data points} $x_i$
at times $t_i$ ($i \in \mathbb{N}$), and it \emph{transmits} a stream of data points $\xi_i$ 
at times $\tau_i$ ($i \in \mathbb{N}$). To simplify the notations, we assume that:
\begin{equation*}
\forall k \in \mathbb{N}, \  \exists ! j \in \mathbb{N} \  \xi_k = t_j
\end{equation*}
That is, transmission times coincide with reception times. 
We define the \emph{centered received points} as follows:
\begin{equation*}
\forall k \in \mathbb{N}\ , \forall i \in \mathbb{N},\ (u^k_i, y^k_i) = (t_{j+i} - t_j, x_{j+i}), \quad \mathrm{where}\ t_j = \xi_k.
\end{equation*}
This definition is such that $u^k_0=0$, and $y^k_i$ is the $i^{th}$ point received
after the $k^{th}$ transmission.

\subsection{LTC algorithm}

Using the notations defined previously, the original LTC algorithm is written as 
follows:
% define x_i (1D), (t, x_i)
\begin{enumerate}
\item Initialization, after $k^{th}$ transmission:
    \begin{itemize}
        \item Wait until first point $(u^k_1, y^k_1)$ is received.
        \item UL $\leftarrow$ $(u^k_1, y^k_1+\epsilon)$
        \item LL $\leftarrow$ $(u^k_1, y^k_1-\epsilon)$
    \end{itemize}
\item Calculate \emph{highLine} as the line connecting $(u^k_0, y^k_0)$ and UL.
\item Calculate \emph{lowLine} as the line connecting $(u^k_0, y^k_0)$ and LL.
\end{enumerate}

@@

\subsection{General formulation}

In this section we provide a distance-independent formulation of LTC in 
dimension $n$. The formulation of LTC in~\cite{ltc} relies on the intersection
of \emph{convex cones} in dimension $n+1$.

When $n$ is higher than 1, the intersection of convex cones unwieldy to 
formalize. Instead, we formulate LTC as an intersection test 
between $m$ \emph{balls} of dimension 
$n$. We formulate LTC recursively, as follows.

Let $x_0 \in \mathbb{R}^n$ be the last data point transmitted by the 
algorithm. For convenience, all the subsequent data points will be 
expressed in the orthogonal space with origin $x_0$. Without loss of 
generality, we assume that the data points are sampled uniformly with a 
sampling period of 1 time unit. Let k be the number of data points 
following $x_0$ that have been compressed.

The intersection of convex cones in dimension $n+1$ can be written as follows:
\ldots

Which is equivalent to the following formulation in $\mathbb{R}^n$:
\begin{eqnarray*}
&& \exists \textbf{u} \in \mathbb{R}^{n} \  s.t. \  \forall i \in \llbracket 1, k \rrbracket, \quad
\norm{i\textbf{u} - \textbf{x}_i} \leq \epsilon\\
\Leftrightarrow && \exists \textbf{u} \in \mathbb{R}^{n} \  s.t. \  \forall i \in \llbracket 1, k \rrbracket, \quad
\norm{\textbf{u} - \frac{1}{i}\textbf{x}_i} \leq \frac{\epsilon}{i} \\
\Leftrightarrow && \bigcap_{i=1}^{k} {\mathcal{B}}_i \neq \O,
\end{eqnarray*}
where $\mathcal{B}_i 
\subset \mathbb{R}^n$ is the ball of center $\frac{1}{i}\textbf{x}_i$ 
and radius $\frac{\epsilon}{i}$.

@@

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,biblio.bib}

>>>>

% packages
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
%\renewcommand{\algorithmicrequire}{\textbf{Input:}}
%\renewcommand{\algorithmicensure}{\textbf{Output:}}
%\usepackage{algpseudocode}
\usepackage{amsfonts} % for R symbol (the set of real numbers)
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[bookmarks=false]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,filecolor=black,urlcolor=blue}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{subcaption}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amsthm}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\title{A multi-dimensional extension of the Lightweight Temporal Compression method}

@@

The algorithm \emph{receives} a stream of \emph{data points} $x_i$
at times $t_i$ ($i \in \mathbb{N}$), and it \emph{transmits} a stream of data points $\xi_i$ 
at times $\tau_i$ ($i \in \mathbb{N}$). To simplify the notations, we assume that:
\begin{equation*}
\forall k \in \mathbb{N}, \  \exists ! i \in \mathbb{N} \  \tau_k = t_i
\end{equation*}
That is, transmission times coincide with reception times. 
We define the \emph{centered received points} as follows:
\begin{equation*}
\forall k \in \mathbb{N}\ , \forall j \in \mathbb{N^*},\ (u^k_j, y^k_j) = (t_{i+j}, x_{i+j}), \quad \mathrm{where\ }i\mathrm{\ is\ s.t.}\ t_i = \tau_k.
\end{equation*}
and:
\begin{equation*}
\forall k \in \mathbb{N},\  (u^k_0, y^k_0) = (\tau_k, \xi_k).
\end{equation*}
This definition is such that $u^k_0=0$, and $y^k_j$ is the $j^{th}$ data point received
after the $k^{th}$ transmission.

Figure~\ref{fig:ltc} illustrates the notations and algorithm.  
\begin{figure}
\includegraphics[width=\columnwidth]{ltc.pdf}
\caption{Illustration of the LTC algorithm and notations used. Red dots indicate transmitted points. Dashed lines
represent the high line and the low line before a point is transmitted.}
\label{fig:ltc}
\end{figure}
\subsection{LTC algorithm}

The LTC algorithm maintains two lines, the \emph{high line}, and the 
\emph{low line} defined by (1) the latest transmitted point and (2) the 
\emph{high point} (high line) and the \emph{low point} (low line). When 
a point ($t_i$, $x_i$) is received, the high line is updated as 
follows: if $x_i+\epsilon$ is below the high line, the high line is 
updated to the line defined by the last transmitted point and ($t_i$, 
$x_i$); otherwise, the high line is not updated. Likewise, the low line 
is updated from $x_i-\epsilon$. Threfore, any line located between the 
high line and the low line approximates the data points received since 
the last transmitted point with an error bounded by $\epsilon$.

Using the notations defined previously, the original LTC algorithm can 
be written as in Algorithm~\ref{algo:ltc}. For readability, we assume 
that access to data points is blocking, i.e., the program will wait 
until the points are available. We also assume that the content of 
variable \texttt{tr} is transmitted after each assignment of this 
variable. Function \texttt{line}, omitted for brevity, returns the 
ordinate at abscissa $x$ (1st argument) of the line defined by the points
in its 2nd and 3rd arguments.

\begin{algorithm}
\begin{algorithmic}[1]
\State tr = $(u^0_0, y^0_0)$ \Comment{last transmitted point}
\State k = 1 ; j = 1
\State lp = $y^1_1 - \epsilon$ \Comment{low point}
\State hp = $y^1_1 + \epsilon$ \Comment{high point}
%\Function{line}{$x$, ($x_0$, $y_0$), ($x_1$, $y_1$)} 
% \Comment{Return the ordinate of the line defined by $(x_0, y_0)$, $(x_1, y_1)$ at abscissa $x$.}
%\Return $y_0 + \frac{y_1-y_0}{x_1-x_0}(x-x_0)$
%\EndFunction
\While{True} \Comment{Process received points as they come}
    \State j += 1
    \State new\_lp = max($y^k_j-\epsilon$, line($u^k_j$, tr, ($u^k_{j-1}$, lp)))
    \State new\_hp = max($y^k_j+\epsilon$, line($u^k_j$, tr, ($u^k_{j-1}$, hp)))
    \If{new\_lp $<$ new\_hp} \Comment{Keep compressing}
        \State lp = new\_lp
        \State hp = new\_hp
        \State \textbf{continue}
    \EndIf
    \State tr = $(u^k_j, (lp+hp)/2)$
    \Comment{Transmit point}
    \State lp = $y^k_j-\epsilon$
    \State hp = $y^k_j+\epsilon$
    \State k += 1
    \State j = 0
\EndWhile
\end{algorithmic}
\caption{Original LTC algorithm, adapted from~\cite{schoellhammer2004lightweight}.}
\label{algo:ltc}
\end{algorithm}

@@

In this section we provide a distance-independent formulation of LTC in 
dimension $n$. By $n$, we refer to the dimension of the data points 
($x_i$). To handle time, LTC actually operates in dimension 
$n+1$.

\subsection{Preliminary comments}

The formulation of LTC in~\cite{schoellhammer2004lightweight} relies on 
the intersection of \emph{convex cones} in dimension $n+1$. For $n=1$, it 
corresponds to the intersection of triangles, which can efficiently be 
computed by maintaining boundary lines, as detailed in the previous 
section. In higher dimension, however, cone intersections are not so 
straighforward, due to the fact that the intersection between 2 cones 
may not be a cone.

To address this issue, we formulate LTC as an intersection test between 
\emph{balls} of dimension $n$, that is, segments for $n=1$, disks for 
$n=2$, etc. Balls are defined with respect to the \emph{norm} used in 
the vector space of data points. For $n=1$, the choice of the norm does 
not really matter, as all p-norms and the infinity norm are identical. 
In dimension $n$, however, norm selection will be critical.

\subsection{Algebraic formulation of LTC}

Let $(u_0^k, y_0^k) \in \mathbb{R}^{n+1}$ be the latest transmitted point. For convenience, all the subsequent points will be 
expressed in the orthogonal space with origin $(u_0^k, y_0^k)$. We denote by $(v_j, z_j)_{j \in \llbracket 0, m \rrbracket}$ such points:
\begin{equation*}
\forall j \leq m,\  (v_j, z_j) = (u_j^k - u_0^k, y_j^k - y_0^k)
\end{equation*}
Without loss of 
generality, we assume that the data points are sampled uniformly with a 
sampling period of 1 time unit.

We define $\mathcal{B}_j$ as the ball of $\mathbb{R}^n$ of center $z_j$ and of radius 
$\epsilon$:
\begin{equation*}
\mathcal{B}_j = \left\{ z \in \mathbb{R}^n,\  \norm{z-z_j} \leq \epsilon \right\} \ (j \leq m)
\end{equation*}
We define $\mathcal{B}'_j$ by plunging $\mathcal{B}_j$ in $\mathbb{R}^{n+1}$:
\begin{equation*}
\mathcal{B'}_j = \left\{ (a_0, \ldots, a_{n}) \in \mathbb{R}^{n+1},\  a_0 = v_j^k \mathrm{\ and\ } (a_1, \ldots, a_{n}) \in \mathcal{B}_j \right\}
\end{equation*}
$\mathcal{B'}_j$ defines a convex cone $\mathcal{C}_j$ in $\mathbb{R}^{n+1}$:
\begin{equation*}
\mathcal{C}_j = \left\{ \alpha x  \in \mathbb{R}^{n+1},\ x \in \mathcal{B'}_j,\ \alpha \in \mathbb{R}^+ \right\}
\end{equation*}

No data point will be transmitted as long as:
\begin{equation*}
\bigcap_{j=0}^m{\mathcal{C}_j} \neq \O
\end{equation*}
Which, due to the definition of convex cones, implies that:
\begin{equation*}
\exists x=(a_0, \ldots a_n) \in \mathbb{R}^{n+1},\ \forall j \in \llbracket 0, m \rrbracket,\ \forall \alpha \in \mathbb{R^+}, \alpha x \in \mathcal{C}_j
\end{equation*}
Without loss of generality, we can assume that $a_0=v^k_1$ since any multiple
of $x$ is also in $\mathcal{C}_j$. 
This implies that $x \in \mathcal{B}'_1$.

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,biblio.bib}
\end{document}

@@
